An assessment of statistical guidance for authors at influential academic journals across scientific disciplines: Preliminary analysis

### 2021-09-01 analysis/preliminaryAnalysis.Rmd
removing gtsummary

```{r load_packages}
library(knitr) # for literate programming
library(papaja) # for article template
library(tidyverse) # for data munging
library(tidylog) # inline analysis feedback
library(here) # for finding files
library(gtsummary) # for summary tables
library(lemon) # to have axes across all graph facets
library(ggthemr) # for ggplot theme
library(ggrepel) # for ggplot text labels
library(scales) # to wrap text on axis labels
```

@@

```{r set_defaults}
theme_gtsummary_compact() # set table theme
ggthemr('fresh') # set ggplot theme
```

@@

**Note:** When journals referred to guidance in external sources, we have recorded the names of the sources, but *we have not currently extracted guidance from those sources*. We return to this point in the *issues to address* section (Section \@ref(issues)) at the end of this document.

@@

**Note:** We found that some publishers provided statistical guidance that was shared across their portfolio of journals (specifically, `r publisher_nature_n` Nature journals, `r publisher_cell_n` Cell journals, and `r publisher_frontiers_n` Frontiers journals). Shared publisher guidance was coded by two team members (as above), but is represented in our data and analysis multiple times - once for each individual journal it applies to. For example, the Cell journals guidance is inherited by `r publisher_cell_n` journals so appears `r publisher_cell_n` times in the dataset. Unless stated otherwise, publisher guidance is treated as journal-specific guidance (i.e., not external guidance) in the analyses presented below.

@@

```{r summarise_data}

# couldn't figure out to extract summary data from gtsummary::tblsummary so making my own summary dataframe
sum_tab <- d_all %>%

@@

sum_tab_general %>%
  filter(esi_field != "OVERALL") %>%
  filter(variable %in% c('guidance', 'internal_guidance_section')) %>%
  ggplot(aes(x=percent, y = 1, colour = str_wrap(esi_field_with_abbr,25))) +
  facet_rep_wrap(facets = vars(var_display), ncol = 2, nrow = 1) +
  coord_capped_cart(bottom='both', left='both') + # to break axes in bottom left corner
  geom_point(size = 5, position = position_stack(), alpha = .5) +
  geom_point(size = 6, position = position_stack(), data = sum_tab_general_overall, colour = 'black', shape = 18) +
  geom_text_repel(
    aes(label = ifelse(percent > 0,esi_abbr,''),
        segment.square = T,
        segment.inflect = T), #
    #point.padding = 0.2,
    force_pull   = 0, # do not pull toward data points
    nudge_y      = 15,
    #nudge_x = 5,
    direction    = "x",
    angle        = 90,
    hjust        = 0,
    segment.size = 0.2,
    segment.curvature = -1e-20,
    max.iter = 1e4, max.time = 1,
    max.overlaps = Inf, # ensure all labels are shown even if overlapping
    show.legend = FALSE # do not add legend entry
  ) +
  scale_y_continuous(limits = c(0,15), name = '1 dot per scientific discipline') +
  xlim(0,100) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major.y = element_blank(),
        legend.position = 'bottom') +
  scale_colour_discrete(name = 'Scientific disciplines') +
  guides(colour = guide_legend(nrow = 6, title.position = 'top'))
```

@@

sum_tab_topics %>%
  filter(esi_field != "OVERALL") %>%
  ggplot(aes(x=percent, y = 1, colour = str_wrap(esi_field_with_abbr,25))) +
  facet_rep_wrap(facets = vars(topic_display), ncol = 2) +
  coord_capped_cart(bottom='both', left='both') + # to break axes in bottom left corner
  geom_point(size = 5, position = position_stack(), alpha = .5) +
  geom_point(size = 6, position = position_stack(), data = sum_tab_topics_overall, colour = 'black', shape = 18) +
  geom_text_repel(
    aes(label = ifelse(percent > 0,esi_abbr,''),
        segment.square = T,
        segment.inflect = T), #
    #point.padding = 0.2,
    force_pull   = 0, # do not pull toward data points
    nudge_y      = 25,
    nudge_x = 10,
    direction    = "x",
    angle        = 90,
    hjust        = 0,
    segment.size = 0.2,
    segment.curvature = -1e-20,
    max.iter = 1e4, max.time = 1,
    max.overlaps = Inf, # ensure all labels are shown even if overlapping
    show.legend = FALSE # do not add legend entry
  ) +
  scale_y_continuous(limits = c(0,25), name = '1 dot per scientific discipline') +
  xlim(0,100) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major.y = element_blank(),
        legend.position = 'bottom') +
  scale_colour_discrete(name = 'Scientific disciplines') +
  guides(colour = guide_legend(nrow = 5, title.position = 'top'))
```

@@

```{r tbl_counts_props}
d_all %>%

  select(esi_field, starts_with("has_")) %>% 
  gtsummary::tbl_summary(
    by = esi_field,
      # statistic = list(everything() ~ "{p}% ({n} / {N})")
  ) %>%
  add_overall %>% 
  modify_footnote(everything() ~ NA) %>% 
  modify_caption("**Number and percentage of journals overall and by scientific discipline providing internal statistical guidance by topic.** Denominator for all proportions is the number of journals in the field, regardless of whether the journal has internal guidance.")
```

@@

```{r phase_two}
# export data for phase two
d_all %>% filter(has_p_value == T | has_significance == T) %>% 
  mutate(coder = "TEH") %>% 
  select(journal, esi_field, p_value, significance) %>%
  mutate(recommended_required = NA,
         p_how = NA,
         p_threshold = NA,
         stat_sig_ok = NA,
         stat_sig_alpha = NA) %>%
  write_csv(here('data','raw','templates','phase_two_pval.csv'))

```

>>>>

```{r load_packages}
library(knitr) # for literate programming
library(papaja) # for article template
library(tidyverse) # for data munging
library(tidylog) # inline analysis feedback
library(here) # for finding files
library(lemon) # to have axes across all graph facets
library(ggthemr) # for ggplot theme
library(ggrepel) # for ggplot text labels
library(scales) # to wrap text on axis labels
```

@@

```{r set_defaults}
ggthemr('fresh') # set ggplot theme
```

@@

**Note:** When journals referred to guidance in external sources, we have recorded the names of the sources, but *we have not extracted guidance from those sources*. We return to this point in the *issues to address* section (Section \@ref(issues)) at the end of this document.

@@

**Note:** We found that some publishers provided statistical guidance that was shared across their portfolio of journals (specifically, `r publisher_nature_n` Nature journals, `r publisher_cell_n` Cell journals, and `r publisher_frontiers_n` Frontiers journals). Shared publisher guidance was coded by two team members (as above), but is represented in our data and analysis multiple times - once for each individual journal it applies to. For example, the Cell journals' guidance is inherited by `r publisher_cell_n` journals so appears `r publisher_cell_n` times in the dataset. Unless stated otherwise, publisher guidance is treated as journal-specific guidance (i.e., not external guidance) in the analyses presented below.

@@

```{r summarise_data}
sum_tab <- d_all %>%

@@

sum_tab_general %>%
  filter(esi_field != "OVERALL") %>%
  filter(variable %in% c('guidance', 'internal_guidance_section')) %>%
  ggplot(aes(x=percent, y = 1, colour = str_wrap(esi_field_with_abbr,25))) +
  facet_rep_wrap(facets = vars(var_display), ncol = 2, nrow = 1) +
  coord_capped_cart(bottom='both', left='both') + # to break axes in bottom left corner
  geom_point(size = 5, position = position_stack(), alpha = .5) +
  geom_point(size = 6, position = position_stack(), data = sum_tab_general_overall, shape = 18, colour = 'black', aes(fill = 'OVERALL')) +
  geom_text_repel(
    aes(label = ifelse(percent > 0,esi_abbr,''),
        segment.square = T,
        segment.inflect = T), #
    #point.padding = 0.2,
    force_pull   = 0, # do not pull toward data points
    nudge_y      = 15,
    #nudge_x = 5,
    direction    = "x",
    angle        = 90,
    hjust        = 0,
    segment.size = 0.2,
    segment.curvature = -1e-20,
    max.iter = 1e4, max.time = 1,
    max.overlaps = Inf, # ensure all labels are shown even if overlapping
    show.legend = FALSE # do not add legend entry
  ) +
  scale_y_continuous(limits = c(0,15), name = '1 dot per scientific discipline') +
  xlim(0,100) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major.y = element_blank(),
        legend.position = 'bottom') +
  guides(colour = guide_legend(nrow = 8, title.position = 'top', title = '')) +
  guides(fill = guide_legend(nrow = 1, title.position = 'top', title = "Scientific disciplines"))
```

@@

sum_tab_topics %>%
  filter(esi_field != "OVERALL") %>%
  ggplot(aes(x=percent, y = 1, colour = str_wrap(esi_field_with_abbr,25))) +
  facet_rep_wrap(facets = vars(topic_display), ncol = 2) +
  coord_capped_cart(bottom='both', left='both') + # to break axes in bottom left corner
  geom_point(size = 5, position = position_stack(), alpha = .5) +
  geom_point(size = 6, position = position_stack(), data = sum_tab_topics_overall, shape = 18, colour = 'black', aes(fill = 'OVERALL')) +
  geom_text_repel(
    aes(label = ifelse(percent > 0,esi_abbr,''),
        segment.square = T,
        segment.inflect = T), #
    #point.padding = 0.2,
    force_pull   = 0, # do not pull toward data points
    nudge_y      = 25,
    nudge_x = 10,
    direction    = "x",
    angle        = 90,
    hjust        = 0,
    segment.size = 0.2,
    segment.curvature = -1e-20,
    max.iter = 1e4, max.time = 1,
    max.overlaps = Inf, # ensure all labels are shown even if overlapping
    show.legend = FALSE # do not add legend entry
  ) +
  scale_y_continuous(limits = c(0,25), name = '1 dot per scientific discipline') +
  xlim(0,100) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major.y = element_blank(),
        legend.position = 'bottom') +
  guides(colour = guide_legend(order = 1, nrow = 8, title.position = 'top', title = '')) +
  guides(fill = guide_legend(order = -1, nrow = 1, title.position = 'top', title = "Scientific disciplines"))
```

@@

```{r tbl_counts_props}
sum_tab %>%
  mutate(field = paste0(esi_field,' N = ',N),
         table_data = paste0(n,' (',percent,'%)')) %>%
  select(field,variable,table_data) %>%
  pivot_wider(id_cols = variable, names_from = field, values_from = table_data) %>%
  select(variable, `OVERALL N = 330`, everything()) %>%
  kable(caption = "**Number and percentage of journals overall and by scientific discipline providing internal statistical guidance by topic.** Denominator for all proportions is the number of journals in the field, regardless of whether the journal has internal guidance.")
```

@@

```{r phase_two}
# export data for phase two
d_all %>% filter(has_p_value == T | has_significance == T) %>% 
  mutate(coder = "TEH") %>% 
  select(journal, esi_field, p_value, significance) %>%
  mutate(recommended_required = NA,
         p_use = NA,
         p_reporting = NA,
         stat_sig_use = NA,
         stat_sig_alpha = NA) %>%
  write_csv(here('data','raw','templates','phase_two_pval.csv'))
```
